### 2025-04-17 sports-cognitive-psychology.qmd
20250408 - placeholder
20250417 - flow

>>>>

Athletes sometimes describe periods of optimal performance where they are "in the zone."
These periods of mental state are called "flow", and are characterized by the following subjective characteristics [@Nakamura2021]:

- intense and focused concentration on the present moment
- merging of action and awareness
- loss of reflective self-consciousness (i.e., loss of awareness of oneself as a social actor)
- a sense that one can control one's actions and deal with the situation because one knows how to respond to whatever happens next
- distorted sense of time (often, a sense that time has passed faster than usual)
- experiencing the activity as intrinsically rewarding

Because these periods often coincide with strong performance, athletes commonly strive to find ways to achieve flow.
